On the Functionality of the Multilayer Perceptron

### 2020-12-14 Thesis/Drafts/Buell-PHYS799-ThesisDraft0.tex
Updates to Thesis

\paragraph*{}The ability to classify time-series data into unique categories has become topic of much attention in the last few decades \cite{Khan,Mierswa}. From climate measurements, to financial reports, to audio-visual media, or to medicine, time-domain signals surround the modern world at every corner \cite{Liu,Zhang}. In each context, the general idea is to place the data or a subset of the data into distinct categories called \textit{classes} based on it's properties. For example, doctors may wish to classify a patient as having a heart condition based on properties of their heartbeat or banks may wish to detect fraudulent transactions based on a history of spending habits. Often, the volume  of information or complexity of the task makes it unreasonable for a human to do unaided by some form of automation.

@@

\paragraph*{}In this work, we have developed an automated neural network program that can classify digital sound waves based on the musical instrument that produced them. The general use of machine learning algorithms for audio information classification or identification is well explored, and a topic of much modern research \cite{Khan,Liu,Mierswa,Serizel,Zhang}.

\paragraph*{}

@@

\paragraph*{}Consider the biological process of hearing a sound wave and matching it to a source. We can model this behavior by some unknown function $F$ and approximate it with the function $F^*$. This can map the contents of a sound file to a potential source as to mimic $F$. For a set of inputs (called features or predictors) $\vec{x} = \big\{ x_0 , x_1 , x_2 , ... , x_{p-1} \big\}$ and a set of classes $0$ through $k-1$, we denote this function as:
\begin{equation}
\label{eqn-MappingFunction}
F^*: \vec{x} \rightarrow \big\{ 0 , 1 , 2 , ... , k-1 \big\}
\end{equation}

\paragraph*{}In secs. (\ref{subsec-NerualNetworkIntro}) and (\ref{subsec-NetworkStructure}), we introduce the nature and structure of a neural network and how it can use used in part to solve this problem. We discuss it's inspiration from a biological brain, and show how different layers of the neural network compare to functions or processes in the brain.

@@

\paragraph*{}In order to train the Multimodal neural network to identify musical instrument sources, we need a suitable data set to present to the model as training data. University of Iowa Electronic Music Studio, and the London-Based Philharmonia Orchestra each have a large collection of publicly available audio files \textcolor{blue}{Citations!}. These files contain short segments of musical instruments performing a single note, and is labeled according to the instrument that produced the sound. The files and labels make up the core of the data set that we use to train the model.

\paragraph*{}To ensure that these data sets are formatted consistently, we read each sample from it's original format, \textit{.AIF}, \textit{.MP3}, or similar, and rewrite the data as a new \textit{.WAV} files, sampled at 44.1 kHz with a 16 bit depth. This ensures that all data will have a consistent format when features are extracted. This stage is detailed in sec. (\ref{subsubsec-Preprocessing}).

@@

\paragraph*{}The performance of a neural network is largely dependent of designing an appropriate set of classification features which represent the input given to the neural network. These are properties of wave forms that can be represented by a numerical value or several numerical values and are used as the primary tool in classification. We detail these features in sec. (\ref{sec-Features}). These are used in place of a full waveform to represent a file's contents in a low-dimensional array. We use tools from physics, mathematics, and signal processing to define and explore a comprehensive set of features that enables a high performance of the classifier.

@@

\paragraph*{}With the appropriate set of features designed, and the number of outputs determined, we can organize the structure of the neural network function. The shape of the network defines the hypothesis space, and a subset of the possible solutions that the trained model could arrive at. We construct a multimodal neural network, explored in sec. (\ref{subsec-Architecture}), that processes two distinct entry points derived from the same audio sample. This network is designed to handle and process each respective input independently, and concatenate the results to produce a single outputted prediction.

@@

\paragraph*{}We divide the raw data set up into a training and testing sets. We employ a resampling method called cross-validation and compute the results of various performance metrics to ensure that our model is making reliable predictions, and can generalize appropriately. In this stage, we also choose the value of hyper-parameters, activation functions and layer widths to best compliment the chosen features. This process is repeated and expanded upon until we have produced a model with a sufficient performance. We explore the explore this is sec. (\ref{sec-PerfEval}).

@@

\paragraph*{}The task of algorithmically matching sound files to musical instrument classes in exceedingly difficult through conventional computer programming techniques. Because of the complexity of sound recognition, the challenge arises as to how to build some sort of program that could function at a level above general procedural or explicit instructional rules. Rather than \textit{hard-coding} a set of conditions or parameters for a classification program, we seek an solution that allows for a computer to \textit{learn} from labeled training examples in order \textit{teach} itself \cite{Bishop,Mitchell}. One such an algorithm that has been used for this type of task is a \textit{Neural Network} \cite{Geron2,Goodfellow,Levine}.

\paragraph*{}As the name implies, early neural networks were inspired from from the human brain. Former YouTube video classification team lead, and current machine learning consultant, Aurelien Geron writes about the relationship between biological brains and mathematical neural networks \cite{Geron}: 
\begin{quote}
Birds inspired us to fly, burdock plants inspired velcro and nature has inspired many other inventions. It seems only logical, then, to look to the brain's architecture for inspiration on how to build an intelligent machine.
\end{quote}
The result of such an analogy is a computer program that is reminiscent of the a brain in that the program can \textit{learn} in a sense. Where a biological brain uses connections of chemical and electrical exchanges to process input and make decisions, the computational counterpart uses arithmetic and numerical exchanges to similarly process input and make decisions \cite{Geron,Levine,Bishop}.

@@

\paragraph*{}A neural network can then be considered as mathematical function that seeks to produce an approximation $F^*$, of some the usually unknown function $F$. The function $F^*$ uses a set of parameters $\Theta$, to map a series of inputs, $x$ (called \textit{features}) to a series of outputs, $y$ (called \textit{predictions}) \cite{Goodfellow,James,Virtanen}. Over the course of their development, humans will be able to learn to match sounds to sources with relative ease, given the appropriate labels. In a similar manner, out neural network is present a collection of sound files and predictors and a complementary set of labels indicating the source of each sound.

\paragraph*{}For this particular project, a neural network has been constructed to perform a \textit{classification} task. A classification task involves mapping the input data, $x$ to one of $k$ possible \textit{classes}, each represented by an integer, $0$ through $k-1$. Each of the $k$ classes represents a particular musical instrument or sound source that could have produced the sound-wave in the audio file \cite{Goodfellow,Loy,Virtanen}. In this section we detail the nature of a general neural network.

@@

\paragraph*{}A Neural Network is a model of a mathematical function, composed of several smaller mathematical functions called \textit{layers} \cite{Goodfellow,Loy}. Each layer represents an operation that takes some real input, typically an array of real double-precision floating-point numbers, and returns a modified array of new double-precision floating-point numbers. The exact nature of this operation can be very different depending on the layer type or where it sits within the network. The output of one layer is passed forward and used as the input for the next layer. It is this process of transforming inputs successively in a particular order until an output is attained that makes up the core functionality of a neural network \cite{Geron,Loy}. The output of the final layer encodes the "decision" of the model given a particular input.

\paragraph*{}Other than inspiration from the brain, Neural Networks are named \textit{networks} due to their successive chained or nested composition nature. A standard representation of a neural network is often a linked list or computational graph structure. This maps out exactly how the repeated compositions are structured and organizes the flow of information in more complicated networks. Each node in the computational graph represents a \textit{layer}, which executes a particular prescribed mathematical function \cite{Goodfellow}. In the case of a \textit{linear, feed-forward} network, information passes in a single direction, to produce an output. We can represent this as a graph:
\begin{equation}
\label{eqn-FunctionGraph}
F^*(x) = x \rightarrow f^{(0)} \rightarrow f^{(1)} \rightarrow ..... \rightarrow
f^{(L-2)} \rightarrow f^{(L-1)} \rightarrow y
\end{equation}
or as a nested function:
\begin{equation}
\label{eqn-FunctionChain}
F(x) = f^{(L-1)}\big[f^{(L-2)}\big[...f^{(1)}\big[f^{(0)}\big[x]\big]\big]...\big]\big]
\end{equation}
The number of layers in a neural network is referred to as the network \textit{depth}. The dimensionality of each layer is referred to as the network \textit{width} \cite{Geron,Loy}.
A network model that contains $L$ unique layers is said to be an $L$-Layer Neural Network, with each layer indexed by a superscipt, $(0)$ through $(L-1)$. Layer $(0)$ is said to be the \textit{input layer} and layer $(L-1)$ is said to be the \textit{output layer} \cite{Geron,Loy}.

@@

\paragraph*{}As stated previously, a neural network is composed of a series of functions that are called in succession to transform features (an input) into predictions (an output). As shown in Eq. (\ref{eqn-altLayerFunction}), each function feeds directly into the next as to form a chain-like computational graph \cite{Goodfellow}. Many layer functions can be divided into two portions: (i.) a Linear transformation, with a bias addition, and (ii.) an element-wise activation transformation. This activation function typically is a non-linear function which allows for the modeling of increasingly complex decision-boundaries \cite{Geron,Loy}. The Linear transformation can be given by Eq.(\ref{eqn-LinearTransform}) in the case of a 1D Dense layer (\ref{subsubsec-DenseLayer}) or Eq. (\ref{eqn-LinearTransform2}) in the case of a 2D Convolution layer.
\begin{equation}
\label{eqn-LinearTransform}
z^{(l)} = W^{(l)} x^{(l-1)} + b^{(l)}
\end{equation}
\begin{equation}
\label{eqn-LinearTransform2}
z^{(l)} = \big(W^{(l)} * x^{(l-1)}\big) + b^{(l)}
\end{equation} 
Where $W^{(l)} x^{(l-1)}$ denotes a matrix-vector product, and $\big(W^{(l)} * x^{(l-1)}\big)$ denotes a convolution product. In both cases, $W^{(l)}$ is the \textit{weighting-matrix} or \textit{weighting-kernel} for layer $l$, $b^{(l)}$ is the \textit{bias-vector} for layer $l$, $z^{(l)}$ are the \textit{linear-activations} for layer $l$ and $x^{(l-1)}$ is the final activations for layer $l-1$.

\paragraph*{}Step (ii.) is usually given by some \textit{activation function}:
\begin{equation}
\label{eqn-elementActivation}
x^{(l)} = \sigma^{(l)} \big[ z^{(l)} \big]
\end{equation}
Where $x^{(l)}$ is the final activations for layer $l$ and $z^{(l)}$ is given in equation (\ref{eqn-LinearTransform}). $\sigma^{(l)}$ is some activation function which introduces the ability to account for non-linearity in the decision boundaries. The combination of Eq. (\ref{eqn-LinearTransform}) and Eq. (\ref{eqn-elementActivation}), for a layer $l$ , make up the function represented by that layer, as in Eq. (\ref{eqn-altLayerFunction}). In this, we discuss and describe the types of layer functions that are used to produce the classification model in this project.

@@

\paragraph*{}The Linear Dense Layer, often just called a \textit{Dense} Layer, or \textit{Fully-Connected} layer, was one of the earliest and most common function types used in artificial neural network models \cite{Geron2,Loy,McCulloch}. A dense layer is composed of a layer of \textit{artificial neurons}, each of which holds a numerical value within it, (usually a double-precision floating point number) called the \textit{activation} of that neuron. This idea was developed from McCulloch and Pitts' work \cite{McCulloch} in attempting to model cognitive processes as mathematical functions, and was expanded upon by Frank Rosenblatt in  1957 by combining many dense layers together to form a multilayer perceptron (MLP) \cite{Geron,Loy,Levine}.

\paragraph*{}We model a layer of neurons as a vector of floating-point numbers. By convention, the array is an $n$-dimensional column vector, where $n$ is number of nodes in that layer. Suppose a layer $(l)$ contains $n$ artificial neurons- we denote the array that hold those activations as $x^{(l)}$ and is given by:
\begin{equation}
\label{layer-DenseNeurons}
\vec{x}^{(l)} = \Big[ x_0, x_1, x_2, \hdots , x_{n-2}, x_{n-1} \Big]^T
\end{equation}
The activation of each entry is given by a linear-combination of activations from the previous layer, Eq.(\ref{eqn-LinearTransform}) and the transformation in Eq.(\ref{eqn-elementActivation}).

@@

\paragraph*{}Convolution layers emerged from studying the brain's visual cortex, and have been used from image recognition related tasks for around 40 years. \cite{Geron,Loy}. The layer function get's it's name from it's implementation of the mathematical \textit{convolution} operation. The 2D discrete-convolution of function or 2D array $A[i,j]$ and $B[i,j]$ is given by \cite{Goodfellow}:
\begin{equation}
\label{eqn-convolution}
C[i,j] = (A * B)[i,j] = \sum_{u}\sum_{v} A[i,j]B[i - u,j - v]
\end{equation}
Note that convolution is commutative: $(A * B) = (B * A)$. We implement a convolutional layer $f^{(l)}$ by creating a series of $K$ weighting matrices, called \textit{filters}, each of size $m \times n$, where $m,n \in \mathbb{Z}$. Often we choose $m = n$ and call the shape the \textit{convolution kernel size} \cite{Loy, Goodfellow}.

\paragraph*{}In the case of a 2D input array, $x^{(l-1)} \in \mathbb{R}^{(M \times N)}$, the convolutions filters "step" through the input data and repeatedly compute the element-wise product of each $m \times n$ weighting kernel, and the local activations of the $x^{(l-1)}$ array. The step size in each of the 2-dimension is known as the \textit{stride}. Each of the $K$ filters are used to generate a $m \times n$ feature map from the input activation. For an appropriately trained network, some filters may result in detecting vertical lines, horizontal lines, sharp edges, areas of mostly one color, etc. \cite{Geron,Loy}.

\paragraph*{}In general, for an $N \times M$ input, with kernel size $m \times n$, with stride size $1 \times 1$, and $K$ feature maps.
\begin{equation}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(M \times N)} \rightarrow
x^{(l)} \in \mathbb{R}^{([M-m+1] \times [N-n+1] \times K)}
\end{equation}
As an example, consider a single filter map over input $x \in \mathbb{R}^{(3 \times 4)}$ and filter map $W \in \mathbb{R}^{(2 \times 2)}$ as shown in Fig. (\ref{fig-2DConvExample}):
\begin{figure}[H]
\begin{center}

	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(a)}
	\caption{Input Activations, $x^{(l-1)}$}
	\end{subfigure}

	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(b)}
	\caption{Convolution Filter, $W^{(l)}_k$}
	\end{subfigure}

	\begin{subfigure}{0.9\textwidth}
	\centering
	\includegraphics[scale=0.6]{../figures/2DConvExample(c)}
	\caption{Convolution Result, $x^{(l)}_k$}
	\end{subfigure}

\end{center}
\caption{The result of convolving an input (a) with an filter map (b) is a new set of activations (c). This Image was adapted from Goodfellow, pg. 325 \cite{Goodfellow}}
\label{fig-2DConvExample}
\end{figure}
Note that formal implementations include a bias vector and an activation function.

\paragraph*{}For 2D convolution over input $x^{(l-1)}$ with feature map $W^{(l)}_k$, producing activations $x^{(l)}$, we compute the activations $x^{(l)}_{k}$ as \cite{Goodfellow}
\begin{equation}
\label{eqn-ConvFeedForward}
x^{(l)}_k(i,j) = \sigma^{(l)}\bigg[ \Big( \sum_{u=0}^{m-1} \sum_{v=0}^{n-1} W^{(l)}_k(i,j) x^{(l-1)}_k(i - u,j - v) \Big) + b^{(l)}_k \bigg]
\end{equation}
This operation repeats for each of the $K$ feature maps. Each maps has it's own $n \times m$ weighting matrix and appropriately shaped bias .

\paragraph*{}The convolution layers allows for several advantages. Among these are (i) \textit{sparse-connectivity}: not every single activation (pixel) is connect to every single output pixel, so it is more computationally efficient, (ii) \textit{positional invariance}: key features can be identified regardless of where they are in the layer, and (iii.) \textit{Automatic feature detection} as the training process will update the filters to identify dominant features in the data without human instruction \cite{Geron,Goodfellow,Loy}. Below we detail what the "Call" method for a 2-dimensional convolutional neural network may look like. Compare this procedure with Eq. (\ref{eqn-ConvFeedForward}).

@@

\FOR {Each feature map, $k \in [0,1,2,...,k-2,k-1]$}
	\item Evaluate linear activation values by computing the 2D convolution-product of 
	$W_{k}$ and $x^{(l-1)}_{k}$:\\
	\item $z_{k}[i,j] \leftarrow b_{k} + 
	\sum_{i=0}^{N-1}\sum_{j=0}^{M-1} W_{k}[u,v] x^{(l-1)}_{k}[u-i,v-j]$ \\
	\item Evaluate output activation values by applying the activation function to each element in the linear activation matrix $z$:\\
	$x^{(l)}_{k}[i,j] \leftarrow \sigma\big[ z_{k}[i,j] \big]$ \\
\ENDFOR

@@

\begin{algorithmic}
\REQUIRE $x^{(l-1)} \leftarrow$ Input activations shaped into $(M \times N)$ matrix
\REQUIRE $m \times n \leftarrow$ Pool height and width
\REQUIRE $P \leftarrow$ A temporary pool array of shape $m \times n $ to hold local activationbs

>>>>

\paragraph*{}The ability to classify time-series data into unique categories has become topic of much attention in the last few decades \cite{Khan,Mierswa}. From climate measurements, to financial reports, to audio-visual media, or to medicine, time-domain signals surround the modern world at every corner \cite{Virtanen}. In each context, the general idea is to place the data or a subset of the data into distinct categories called \textit{classes} based on it's properties. For example, doctors may wish to classify a patient as having a heart condition based on properties of their heartbeat or banks may wish to detect fraudulent transactions based on a history of spending habits. Often, the volume  of information or complexity of the task makes it unreasonable for a human to do unaided by some form of automation.

@@

\paragraph*{}In this work, we have developed an automated neural network program that can classify digital sound waves as one of $37$ musical instrument that it most closely resembles. The general use of machine learning algorithms for audio information classification or identification is well explored, and a topic of much modern research \cite{Khan,Liu,Mierswa,Serizel,Zhang}. From previous work, it has been found that various types of audio classification are possible, and are still being expanded upon. Some papers details features and performances of attempting to classify speech against music or organize audio-visual media based on content.

\paragraph*{}It is well know that the choice of inputs, called features is critically important to classification success \cite{Khan,Serizel}. In many cases, we can develop and explore a set of features that allows for the particular classification task to be executed considerably well. In our case, the choice of features is motivated by the physics of each musical instrument class. By exploring the nature of the vibration part of each instrument according to the Hornbostel-Sachs classification system, we can describe some of the mechanical properties and see how they differ between instruments of different types. These differences are key to providing a set of features which allow a neural network to differentiate between classes of instruments \cite{Liu,Zhang}.

\paragraph*{}The design of the neural function that we have constructed combines two architecture types by using two distinct entry arrays which are processed, and a single output is formed. Thus type of neural network is referred to as a \textit{multimodal} nueral network \cite{Li,Ngiam}.

\paragraph*{}\textcolor{red}{This is not my best into to date}

@@

\paragraph*{}Consider the biological process of hearing a sound wave and matching it to a source. We can model this behavior by some unknown function $F$ and approximate it with the function $F^*$. This can map the contents of a sound file to a potential source as to mimic the behavior of $F$. For a set of inputs (called features or predictors) $\vec{x} = \big\{ x_0 , x_1 , x_2 , ... , x_{p-1} \big\}$ and a set of classes $0$ through $k-1$, we denote this function as:
\begin{equation}
\label{eqn-MappingFunction}
F^*: \vec{x} \rightarrow \big\{ 0 , 1 , 2 , ... , k-1 \big\}
\end{equation}

\paragraph*{}In secs. (\ref{subsec-NerualNetworkIntro}) and (\ref{subsec-NetworkStructure}), we introduce the nature and structure of a neural network and how it can use used in part to solve this problem. We discuss it's inspiration from a biological brain,and show how different layer types are influenced from the brain's functions as well as how they are used in this problem.

@@

\paragraph*{}In order to train the neural network to be able to identify musical instrument sources, we need a suitable data set to present to the model as training data. University of Iowa Electronic Music Studio, and the London-Based Philharmonia Orchestra each have a large collection of publicly available audio files (See acknowledgments). These files contain short segments of musical instruments performing a single note, and is labeled according to the instrument that produced the sound. The files and labels make up the core of the data set that we use to train the model.

\paragraph*{}To ensure that these data sets are formatted consistently, we read each sample from it's original format, \textit{.AIF}, \textit{.MP3}, or similar, and rewrite the data as a new \textit{.WAV} files, sampled at 44.1 kHz with a 24-bit depth. This ensures that all data will have a consistent format when features are extracted. This stage is detailed in sec. (\ref{subsubsec-Preprocessing}).

@@

\paragraph*{}The performance of a neural network is largely dependent of designing an appropriate set of predictors or features which represent the raw input given to the model. These inputs are designed to represent properties of wave forms that can be expressed by a single or multiple numerical values. We detail these features in sec. (\ref{sec-Features}). These are used in place of a full waveform to represent a file's contents in a low-dimensional array. We use tools from physics, mathematics, and signal processing to define and explore a comprehensive set of features that enables a high performance of the classifier.

@@

\paragraph*{}With the appropriate set of features designed, and the number of output categories determined, we can organize the structure of the neural network function. The shape of the network defines the flow of information, and the hypothesis space. The nature of all possible solutions that the trained model could arrive at is subject to the architecture \cite{Goodfellow,Loy}. We construct a multimodal neural network, explored in sec. (\ref{subsec-Architecture}), that processes two distinct entry points derived from the same audio sample. This network is designed to evaluate each input independently, and concatenate the results to produce a single outputted prediction.

@@

\paragraph*{}We execute a K-Folds cross validation algorithm to evaluate the performance of the neural network across overlapping subsets of data. This ensures that the model can perform consistently and reliably given small changes in initial conditions. In this stage, we also choose the value of hyper-parameters, activation functions and layer widths to best compliment the chosen features. This process is repeated and expanded upon until we have produced a model with a sufficient performance. We explore the explore this is sec. (\ref{sec-PerfEval}).

@@

\paragraph*{}The task of automatically matching sound files to musical instrument classes is exceedingly challenging through conventional computer programming techniques. Because of the complexity of sound recognition, the challenge arises as to how to build some sort of program that could function at a level above general procedural or explicit instructional rules. Rather than \textit{hard-coding} a set of conditions or parameters for a classification program, we seek an solution that allows for a computer to \textit{learn} from labeled training examples in order \textit{teach} itself \cite{Bishop,Mitchell}. We choose to employ a neural network, which a computer algorithm that is designed to update itself to perform a task with increasing proficiency as it is presented with new information \cite{Geron2,Goodfellow,Levine,Mitchell}.

\paragraph*{}As the name implies, early neural networks were inspired from from the human brain. Former YouTube video classification team lead, and current machine learning consultant, Aurelien Geron writes about the relationship between biological brains and mathematical neural networks \cite{Geron}: 
\begin{quote}
Birds inspired us to fly, burdock plants inspired velcro, and nature has inspired many other inventions. It seems only logical then, to look to the brain's architecture for inspiration on how to build an intelligent machine.
\end{quote}
The result is a computer program that is reminiscent of the a brain in that it structured much like the brain, and can\textit{learn} from new data. Where a biological brain uses connections of chemical and electrical exchanges to process input and make decisions, the computational counterpart uses arithmetic and numerical exchanges to similarly process input and make decisions \cite{Levine,Bishop}. Where a brain uses physical connection of neurons through axons, the nueral network uses mathematical connections of neurons through mathematical operations \cite{Geron,Levine}.

@@

\paragraph*{}The neural network function $F^*$ uses a set of parameters $\Theta$, to map those inputs, $x$ to outputs, $y$ \cite{Goodfellow,James,Virtanen}. It is the process of updating the elements in the object $\Theta$ which allow us to train a neural network to perform the desired task. Similarly, it is the process of organizing and connecting neurons in the brain that allow humans to perform any specific task \cite{Geron,Levine}. Over the course of their lives, humans will be able to learn to match sounds to sources with relative ease, given the appropriate labels. In a similar manner, our neural network is presented a collection of predictors from sound files and a complementary set of labels indicating the source of each sound. This type of task is called \textit{classification}, which involves mapping the input data, $x$ to one of $k$ possible \textit{classes}, each represented by an integer, $0$ through $k-1$. Each of the $k$ classes represents a particular musical instrument or sound source that could have produced the sound-wave in the audio file \cite{Goodfellow,Loy,Virtanen}. In this section, we explore the motivation, structure, and training process of a neural network.

@@

\paragraph*{}A Neural Network is a model of a mathematical function, composed of several smaller mathematical functions called \textit{layers} \cite{Goodfellow,Loy}. Each layer represents an operation that takes an input, typically an array of real double-precision floating-point numbers, and returns an output which is a modified array of new double-precision floating-point numbers. The exact nature of this operation can be very different depending on the layer type or where it sits within the network. Typically, a neural network is characterized by a chain-like structure where the output of one layer is fed directly into the input of the next layer \cite{Geron,Goodfellow}. It is this process of transforming inputs successively in a particular order until an output is attained that makes up the core functionality of a neural network \cite{Geron,Loy}. The output of the final layer encodes the "decision" of the model given a particular input.

\paragraph*{}Other than inspiration from the brain, Neural Networks are named \textit{networks} due to their successive chained or nested composition nature. A standard representation of a neural network is often a linked list or computational graph structure \cite{Goodfellow}. This maps out exactly how the repeated compositions are structured and organizes the flow of information in more complicated networks. Each node in the computational graph represents a \textit{layer}, which executes a particular prescribed mathematical function \cite{Goodfellow}. In the case of a \textit{linear, feed-forward} network, such as the one that we implement, information passes in a single direction to produce an output. We can represent this as a graph:
\begin{equation}
\label{eqn-FunctionGraph}
F^*(x) = x \rightarrow f^{(0)} \rightarrow f^{(1)} \rightarrow ..... \rightarrow
f^{(L-2)} \rightarrow f^{(L-1)} \rightarrow y^*
\end{equation}
or as a nested function:
\begin{equation}
\label{eqn-FunctionChain}
F^*(x) = f^{(L-1)}\big[f^{(L-2)}\big[...f^{(1)}\big[f^{(0)}\big[x\big]\big]\big]...\big]\big] = y^*
\end{equation}
The number of layers in a neural network is referred to as the network \textit{depth}. The dimensionality of each layer is referred to as the network \textit{width} \cite{Geron,Loy}.
A network model that contains $L$ unique layers is said to be an $L$-Layer Neural Network, with each layer indexed by a superscipt, $(0)$ through $(L-1)$. Layer $(0)$ is said to be the \textit{input layer} and layer $(L-1)$ is said to be the \textit{output layer} \cite{Geron,Loy}.

@@

\paragraph*{}As stated previously, a neural network is composed of a series of functions that are called in succession to transform features (an input) into predictions (an output). As shown in Eq. (\ref{eqn-altLayerFunction}), each function feeds directly into the next as to form a chain-like computational graph \cite{Goodfellow}. There are multiple type of layers that we use to construct the network in this project, but the two most important are called a \textit{dense} layer and a \textit{convoltuion} layer. These layers each contain a set of parameters which makeupa  subset of the trainable elements in $\Theta$.
These two layer functions can be divided into two portions: (i.) a Linear transformation, with a bias addition, and (ii.) an element-wise activation transformation. The Linear transformation can be given by Eq.(\ref{eqn-LinearTransform}) in the case of a 1D Dense layer (\ref{subsubsec-DenseLayer}) or Eq. (\ref{eqn-LinearTransform2}) in the case of a 2D Convolution layer.
\begin{equation}
\label{eqn-LinearTransform}
z^{(l)} = W^{(l)} x^{(l-1)} + b^{(l)}
\end{equation}
\begin{equation}
\label{eqn-LinearTransform2}
z^{(l)} = \big(W^{(l)} * x^{(l-1)}\big) + b^{(l)}
\end{equation} 
Where $W^{(l)} x^{(l-1)}$ denotes a matrix-vector product, and $\big(W^{(l)} * x^{(l-1)}\big)$ denotes a convolution product. In both cases, $W^{(l)}$ is a \textit{weighting-matrix} or \textit{weighting-kernel} for layer $(l)$, $b^{(l)}$ is the \textit{bias-vector} for layer $(l)$, $z^{(l)}$ are the \textit{linear-activations} for layer $(l)$ and $x^{(l-1)}$ is the final activations for layer $(l-1)$.

\paragraph*{}Step (ii.) is usually given by some \textit{activation function}:
\begin{equation}
\label{eqn-elementActivation}
x^{(l)} = \sigma^{(l)} \big[ z^{(l)} \big]
\end{equation}
Where $x^{(l)}$ is the final activations for layer $l$ and $z^{(l)}$ is given in equation (\ref{eqn-LinearTransform}). $\sigma^{(l)}$ is some activation function which introduces the ability to account for complicated non-linear decision boundaries which the neural network uses to make predictions. The combination of Eq. (\ref{eqn-LinearTransform}) and Eq. (\ref{eqn-elementActivation}), for a layer $l$ , make up the function represented by that layer, as in Eq. (\ref{eqn-altLayerFunction}). In this section, we detail the types of layer functions that are used to produce the classification model in this project.

@@

\paragraph*{}The Linear Dense Layer, often just called a \textit{Dense} Layer, or \textit{Fully-Connected} layer, was one of the earliest and most common function types used in artificial neural network models \cite{Geron2,Loy,McCulloch}. A dense layer is composed of a layer of \textit{artificial neurons}, each of which holds a numerical value within it, (usually a double-precision floating point number) called the \textit{activation} of that neuron. The activation of a dense layer are given by a linear combination of the layers inputs. This idea was developed from McCulloch and Pitts' work \cite{McCulloch} in attempting to model cognitive processes as mathematical functions by connecting neurons through a series of weighting values. The idea was expanded upon by Frank Rosenblatt in  1957 by combining many dense layers together to form a multilayer perceptron (MLP), which will make up a portion of the network that we have implemented. \cite{Geron,Loy,Levine}.

\paragraph*{}We model a layer of neurons as an array or vector of floating-point numbers. Suppose a layer $(l)$ contains $n$ artificial neurons- we denote the $n$-dimensional array that holds those activations as $x^{(l)}$ and is given by:
\begin{equation}
\label{layer-DenseNeurons}
\vec{x}^{(l)} = \Big[ x_0, x_1, x_2, \hdots , x_{n-2}, x_{n-1} \Big]^T
\end{equation}
The activation of each entry is given by a linear-combination of activations from the previous layer, Eq.(\ref{eqn-LinearTransform}) and the transformation in Eq.(\ref{eqn-elementActivation}).

@@

\paragraph*{}Convolution layers emerged from studying the brain's visual cortex, and have been used from image recognition related tasks for around 40 years. \cite{Geron,Loy}. The layer function get's it's name from it's implementation of the mathematical \textit{convolution} operation. The 2D discrete-convolution of function or 2D array $A[i,j]$ and $B[i,j]$ is given by \cite{Goodfellow}:
\begin{equation}
\label{eqn-convolution}
C[i,j] = (A * B)[i,j] = \sum_{u}\sum_{v} A[i,j]B[i - u,j - v]
\end{equation}
Note that convolution is commutative: $(A * B) = (B * A)$. We implement a convolutional layer $f^{(l)}$ by creating a series of $K$ weighting matrices, called \textit{filters}, each of size $m \times n$, where $m,n \in \mathbb{Z}$. Often we choose $m = n$ and we call the shape the \textit{convolution kernel size} \cite{Loy, Goodfellow}. The step size in each of the two dimensions is known as the \textit{stride size}, which we choose to be $1$ in both dimensions.

\paragraph*{}In the case of a 2D input array, $x^{(l-1)} \in \mathbb{R}^{(M \times N)}$, the convolutions filters "step" through the input data and repeatedly compute the element-wise product of each $m \times n$ weighting kernel, and the local activations of the $x^{(l-1)}$ array. Each of the $K$ filters are used to generate a $m \times n$ feature map from the input activations. For an appropriately trained network, some filters may result in detecting vertical lines, horizontal lines, sharp edges, areas of mostly one color, etc. \cite{Geron,Loy}.

\paragraph*{}In general, for an $M \times N$ input, with kernel size $m \times n$, with stride size $1 \times 1$, and $K$ feature maps.
\begin{equation}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(M \times N)} \rightarrow
x^{(l)} \in \mathbb{R}^{([M-m+1] \times [N-n+1] \times K)}
\end{equation}
As an example, consider a single filter map over input $x \in \mathbb{R}^{(3 \times 4)}$ and filter map $W \in \mathbb{R}^{(2 \times 2)}$ as shown in Fig. (\ref{fig-2DConvExample}):
\begin{figure}[H]
\begin{center}
	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(a)}
	\caption{Input Activations, $x^{(l-1)}$}
	\end{subfigure}	
	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(b)}
	\caption{Convolution Filter, $W^{(l)}_k$}
	\end{subfigure}	
	\begin{subfigure}{0.9\textwidth}
	\centering
	\includegraphics[scale=0.6]{../figures/2DConvExample(c)}
	\caption{Convolution Result, $x^{(l)}_k$}
	\end{subfigure}	
\end{center}
\caption{The result of convolving an input (a) with an filter map (b) is a new set of activations (c). This Image was adapted from Goodfellow, pg. 325 \cite{Goodfellow}}
\label{fig-2DConvExample}
\end{figure}
Note that formal implementations include a bias vector and an activation function.

\paragraph*{}For 2D convolution layer $(l)$, given input $x^{(l-1)}$, we compute the activations of the $k$-th filter $x^{(l)}_{k}$ using feature map $W^{(l)}_k$ and bias neuron $b^{(l)}_k$ as \cite{Goodfellow}:
\begin{equation}
\label{eqn-ConvFeedForward}
x^{(l)}_k(i,j) = \sigma^{(l)}\bigg[ b^{(l)}_k +  \sum_{u=0}^{m-1} \sum_{v=0}^{n-1} W^{(l)}_k(i,j) x^{(l-1)}_k(i - u,j - v) \bigg]
\end{equation}
This operation repeats for each of the $K$ feature maps. Each maps has it's own $n \times m$ weighting matrix and appropriately shaped bias .

\paragraph*{}The convolution layers allows for several advantages over the dense layer. Among these are (i) \textit{sparse-connectivity}: not every single activation (pixel) is connect to every single output pixel, so it is more computationally efficient, (ii) \textit{positional invariance}: key features can be identified regardless of where they are in the layer, and (iii.) \textit{Automatic feature detection} as the training process will update the filters to identify dominant features in the data without human instruction \cite{Geron,Goodfellow,Loy}. Below we detail what the "Call" method for a 2-dimensional convolutional neural network may look like. Compare this procedure with Eq. (\ref{eqn-ConvFeedForward}).

@@

\FOR {Each feature map, $k \in [0,1,2,...,K-2,K-1]$}
	\item Evaluate linear activation values by computing the 2D convolution-product of 
	$W_{k}$ and $x^{(l-1)}_{k}$:\\
	\item $z_{k}[i,j] \leftarrow b_{k} + 
	\sum_{i=0}^{N-1}\sum_{j=0}^{M-1} W_{k}[u,v] x^{(l-1)}_{k}[u-i,v-j]$ \\
	\item Evaluate output activation values by applying the activation function to each element in the linear activation matrix $z$:\\
	$x^{(l)}_{k}[i,j] \leftarrow \sigma\big[ z_{k}[i,j] \big]$ \\
\ENDFOR

@@

\begin{algorithmic}
\REQUIRE $x^{(l-1)} \leftarrow$ Input activations shaped into $(M \times N)$ matrix
\REQUIRE $m \times n \leftarrow$ Pool height and width
\REQUIRE $P \leftarrow$ A temporary pool array of shape $m \times n $ to hold local activations
